---
title: Data  types
author: ''
date: '2022-02-21'
slug: data-types
categories: []
tags: []
---
## Call Libraries 
```{r}
library(tidyverse)
library(rmarkdown)
```

## Import files from Github for cat lover data

```{r}
url1 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/cat-lovers.csv'
cat_lovers <- read_csv(url1)

```

## Running summarise fails due to non-numeric arguments
```{r}
cat_lovers %>%
 summarise(mean_cats = mean(number_of_cats))
```
### Using the help command on mean shows that there is an argument that takes out NA values however including this argument does not fix the error

```{r}
?mean
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats, na.rm = TRUE))
```
## Running the type command shows us we have data type mismatch
```{r}
glimpse(cat_lovers)
```
### We were expecting a numeric data type but we recieved chr instead so we must explicitly convert the data type
```{r}
cat_lovers %>%
 mutate(number_of_cats = case_when(
     name == "Ginger Clark"  ~ 2,
     name == "Doug Bass"     ~ 3,
     TRUE                    ~as.numeric (number_of_cats)
    )) %>%
  summarise(mean_cats = mean(number_of_cats))
```
### Convert data types after manipulating data to the same type to ensure all points are numeric type and data is all one type
```{r}
cat_lovers %>%
 mutate(number_of_cats = case_when(
     name == "Ginger Clark"  ~ "2",
     name == "Doug Bass"     ~ "3",
     TRUE                    ~ number_of_cats
    ),
number_of_cats =  as.numeric(number_of_cats)
) %>%
  summarise(mean_cats = mean(number_of_cats))
```
## Save value as a variable
```{r}
cat_lovers <- cat_lovers %>%
 mutate(number_of_cats = case_when(
     name == "Ginger Clark"  ~ "2",
     name == "Doug Bass"     ~ "3",
     TRUE                    ~ number_of_cats
    ),
number_of_cats =  as.numeric(number_of_cats)
) %>%
  summarise(mean_cats = mean(number_of_cats))
```
