---
title: Data IO
author: ''
date: '2022-02-23'
slug: data-io
categories: []
tags: []
---
## Call Libraries 
```{r}
library(tidyverse)
library(rmarkdown)
library(readr)
library(ggplot2)
library(readxl)
```
## Read the data from the CSV file
```{r}
  nobel <- read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/nobel.csv')
nobel
```
### Split data into two data frames one for STEM and one for non-STEM

```{r}
nobel_stem <- nobel %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
                         
nobel_nonstem <- nobel %>%
  filter(category != "Physics" & category != "Medicine" & category != "Chemistry" & category != "Economics")

```

## Write CSV files to the folder containing this post
```{r}
write_csv(nobel_stem, file = "C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/nobel_stem.csv") 

write_csv(nobel_nonstem, file = "C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/nobel_nonstem.csv")
```

### Readback CSV files to inspect

```{r}
read_csv("C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/nobel_stem.csv")
read_csv("C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/nobel_nonstem.csv")
```
# Naming Data

```{r}
edibnb_badnames <- read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/edibnb-badnames.csv')
names(edibnb_badnames)
```
## R does not allow spaces in variable names so we have some optionns
#### Option 1: Define coloumn names

```{r}
edibnb_col_names <- read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/edibnb-badnames.csv',
        col_names = c("id", "price","neighbourhood","accommodates", "bathroom", "bedroom","bed", "review_scores_rating",                   "n_reviews", "url"))
names(edibnb_col_names)
```
#### Option 2: Format text to snake_case
```{r}
edibnb_clean_names <- read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/edibnb-badnames.csv') %>%
  janitor::clean_names()
names(edibnb_clean_names)
```
# Variable types
### Types shown in output

```{r}
read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/df-na.csv')
```
## Although it is chr we want x to be double so we assign explicit NA
```{r}
read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/df-na.csv',
         na = c("", "NA", "9999", "Not applicable")) 
```
## Otherwise, we can specifcy coloumn types
```{r}
read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/df-na.csv', col_types = list(col_double(),
                                             col_character(),
                                             col_character()))

```
## Case study: Favorite Foods
```{r}
fav_food <- read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/favourite-food.csv')
fav_food
```
## Clean variable names

```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/vsomesula/tidyverse/main/favourite-food.csv") %>%
  janitor::clean_names()
fav_food
```
## Handle the NA's

```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/vsomesula/tidyverse/main/favourite-food.csv", na = c("N/A", "99999")) %>%
  janitor::clean_names()
fav_food
```
## Make age numeric

```{r}
fav_food <- fav_food %>%
  mutate(
    age = if_else(age == "five", "5", age),
    age = as.numeric(age)
    )
glimpse(fav_food)
```

## Check socio-economic status
```{r}
fav_food %>%
  count(ses)
```
### Make ses a factor
```{r}
fav_food <- fav_food %>%
  mutate(ses = fct_relevel(ses, "Low", "Middle", "High"))
fav_food %>%
  count(ses)
```

## Put all the code together for a cleaned up version of your code

```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/vsomesula/tidyverse/main/favourite-food.csv", na = c("N/A", "99999")) %>%
  janitor::clean_names() %>%
  mutate(
    age = if_else(age == "five", "5", age), 
    age = as.numeric(age),
    ses = fct_relevel(ses, "Low", "Middle", "High")
  )
fav_food
```

## Write the CSV file and read it back in to check 

```{r}
write_csv(fav_food, file = "C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/fav-food-clean.csv")
fav_food_clean <- read_csv("C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/fav-food-clean.csv")

```
## Check what happens to ses

```{r}
fav_food_clean %>%
  count(ses)
```
# Sometimes an rds file type is more convinient
```{r}
write_rds(fav_food, file = "C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/fav-food-clean.rds")
fav_food_clean <- read_rds("C:/Users/prana/Documents/COPwebsite/content/post/2022-02-23-data-io/fav-food-clean.rds")
fav_food_clean %>%
  count(ses)
```
## Your turn
```{r}
sales <- read_csv('https://raw.githubusercontent.com/vsomesula/tidyverse/main/sales.csv', col_names = c("Brand", "ID", "N")) %>%
  janitor :: clean_names() %>% 
  mutate(
    brand = if_else(brand == "B1", "Brand 1", "Brand 2")
      )
sales <- sales[-1,]
sales
```



