---
title: Working with your multiple data frames
author: ''
date: '2022-02-16'
slug: working-with-your-multiple-data-frames
categories: []
tags: []
---
## Call Libraries 
```{r}
library(tidyverse)
```

## Import files from Github for top 10 women scientists

```{r}
url1 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/professions.csv'
url2 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/dates.csv'
url3 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/works.csv'
professions <- read_csv(url1)
dates <- read_csv(url2)
works <- read_csv(url3)
```
## Left join dates and professions data frames
```{r}
professions %>%
  left_join(dates)
```
## Right join dates and professions data frames
```{r}
professions %>% 
  right_join(dates)
```

## Full joining the first works and dates dataframes
```{r}
dates %>% 
  full_join(works)
```
## Inner join works and dates dataframes
```{r}
dates %>%
    inner_join(works)
```
## Semi join works and dates dataframes
```{r}
dates %>%
  semi_join(works)
```

## Anti join works and dateframes
```{r}
dates %>%
  anti_join(works)
```

## Combining professions, dates, and works dataframes
```{r}
professions %>%
  left_join(dates) %>%
  left_join(works)
```

## Import dataframes for student records

```{r}
url4 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/enrolment.csv'
url5 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/survey.csv'
enrolment <- read_csv(url4)
survey <- read_csv(url5)

```

## Student records for students in class
```{r}
enrolment %>%
  left_join(survey, by = "id")
```

## Student records for students surveyed missing
```{r}
enrolment %>%
  anti_join(survey, by = "id")
```

## Student records for students dropped

```{r}
survey %>%
  anti_join(enrolment, by = "id")
```

## Import dataframes for grocery sales

```{r}
url6 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/purchases.csv'
url7 <- 'https://raw.githubusercontent.com/vsomesula/tidyverse/main/prices.csv'
purchases <- read_csv(url6)
prices <- read_csv(url7)

```

## Find grocery sales total revenue

```{r}
purchases %>%
    left_join(prices) %>%
    summarise(total_revenue = sum(price))
```

## Find grocery sales revenue per customer

```{r}
purchases %>%
  left_join(prices) %>%
  group_by(customer_id) %>%
  summarise(total_revenue = sum(price))
```

