---
title: 'A02: Regression Models Exercise'
author: ''
date: '2022-03-26'
slug: a02-regression-models-exercise
categories: []
tags: []
---
# Executive Summary
A machine learning professor wants to apply his skills towards developing a model to price diamonds so he can determine if he is getting a fair price for an engagement ring he wants to buy. He determined that the variables used in pricing a diamond are color, cut, carat weight, clarity, polish, symmetry, and certification. He has collected data from three different diamond websites to assist us in developing a multiple regression model to predict the price of a diamond ring of his interest.
INCLUDE LAST LINE OF SUMMARY OF MODEL

# Issues and challenges 
In order to develop a multiple regression model, it is important to have a combination of metric variables and the categorical variables that are ordinally scaled. This means it is important to determine how the data in the categorical variables will be grouped.

# Setup
```{r}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(pacman,readxl, magrittr, productplots, psych, RColorBrewer, tidyverse, boot, caret)
diamonds <- read_csv("https://raw.githubusercontent.com/vsomesula/COPwebsite/main/content/post/2022-03-26-a02-regression-models-exercise/Professor_Proposes_Data.csv")
diamonds
```
```{r,echo=FALSE}
diamonds <- read_excel("C:/Users/prana/Downloads/Professor_Proposes_Data.xlsx")
```
```{r}
attach(diamonds)
diamonds$Colour <- as.factor(Colour)
diamonds$Clarity <- as.factor(Clarity)
diamonds$Cut <- as.factor(Cut)
diamonds$Certification <- as.factor(Certification)
diamonds$Polish <- as.factor(Polish)
diamonds$Symmetry <- as.factor(Symmetry)
diamonds$Wholesaler <- as.factor(Wholesaler)
diamonds
```
# Graphing and frequency counting of each variable
```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Carat, fill =  Carat)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Carat)
summary(Carat)

```
### The data ranges from 0.09 to 1.58 and this a follows biomodal distrubition with two peaks with one around 0.2 and the other around 1.0


```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Colour, fill =  Colour)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Colour)

```
### Colour could be mutated to yellow and colorless as the different ratings are simply a range for both of these colors. One grouping would be D-I as colorless and the rest as yellow


```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Clarity, fill =  Clarity)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Clarity)

```

### Instead of being ordinal data, this data could be converted to nominal data since there is an inherent order by assigning the numbers 1-9 to each "level" in the Clarity variable


```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Cut, fill =  Cut)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Cut)

```

### The data seems to be well balanced


```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Certification, fill =  Certification)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Certification)

```
### The certification value seems to be ordinal data with high representation in the GIA and EGL category.


```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Polish, fill =  Polish)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Polish)

```
### Some categories such as Fair, Ideal, and even Excellent to an extent are not well represented so they can be collapsed into other categories such as Fair-Ideal, and Very Good-Excellentl.

```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Symmetry, fill =  Symmetry)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Symmetry)

```

### For symmetry certain categories such as Ideala nd Fair are less represented and it could similairly be collpased like Polish.


```{r}
diamonds %>%
  ggplot() +
  geom_bar(aes(Wholesaler, fill =  Wholesaler)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
count(diamonds, Wholesaler)

```
### This is obvious ordinal data that cannot be simplified.

# Transformation of Data
```{r}
diamonds <- mutate(diamonds, Colour2C = ifelse(Colour == "J", "Colourless", ifelse(Colour == "K", "Colourless", ifelse(Colour == "L", "Colourless", "Yellow"))))
diamonds<- mutate(diamonds, Polish2C = ifelse(Polish == "v", "VG-E", ifelse(Polish == "V,", "VG-E", ifelse(Polish == "X", "VG-E", "F-I"))))
diamonds<- mutate(diamonds, Symmetry2C = ifelse(Symmetry == "v", "VG-E", ifelse(Symmetry == "V,", "VG-E", ifelse(Symmetry == "X", "VG-E", "F-I"))))
attach(diamonds)
```


# Exploratory Analysis
## This type of model will give us an equation where we can use as many of the categorial variables as we want. We can imagine it as CutGx1 + CutIx2 + .. with their respective coefficents. With each diamond if it has Cut level of G than x2, x3, etc for that category will be 0 leaving us only with the weight for Cut G. This will leave us a very convuluted model however as seen below.
```{r}
lm.fit <- lm(Price~Carat + Colour2C + Clarity +Cut + Certification + Polish2C + Symmetry2C, data = diamonds)
summary(lm.fit)
caret::varImp(lm.fit)
lm.fit.untransformed <- lm(Price~Carat + Colour + Clarity +Cut + Certification + Polish + Symmetry, data = diamonds)
summary(lm.fit.untransformed)
caret::varImp(lm.fit.untransformed)
```
# Analysis of model

There is much to unpack from running the lines of code above. Just for comparison, I ran two models one where the variables had been transformed and one where the variables had not been transformed. I also used the caret package and the variable importance function within the package to determine which variables were most significant - or in other words which variables to include in the final model.

First we can see that Carat is a huge determining variable for price. This is corroborated by the huge number the Carat function returned for both variables. Similarly, the Colour variable in its transformed form is shown to be an important variable due to its low p value. Polish and Symmetry despite being transformed do not seem to be variables with high predictive power. Clarity also seems to be high predictive variables as seen by their low p values and high values in the varImp function. So in conclusion, Carat, Clarity, and Color are the most important variables for determine price. Cut and certification seem to be the middle tier variables for determining price. Finally, polish and symmetry have minimal impact in determining the price

## We can use this model for testing however through code to see the results. 
```{r}
df <- data.frame(Carat = 0.9, Clarity = "SI2", Cut = "V", Certification = "GIA", Colour2C = "Colourless", Polish2C = "F-I", Symmetry2C = "VG-E")
as_tibble(df)
df$Colour2C <- as.factor(df$Colour)
df$Clarity <- as.factor(df$Clarity)
df$Cut <- as.factor(df$Cut)
df$Certification <- as.factor(df$Certification)
df$Polish2C <- as.factor(df$Polish)
df$Symmetry2C <- as.factor(df$Symmetry)
df
predict(lm.fit, df )
```
# The final price the model has generated is 2736.82 which is different from the 3100 generated by the professor. However, we did not consider the labor cost of retail and other associated costs of the diamond such as transport of the diamond that may be factored into the diamonds price. Furthermore, the ring was not included in the data provided.

# Boot strap validation of the model
```{r}
boot.fn <- function(data, index){
  return(coef(lm(Price~Carat+Colour2C+Cut+Polish2C+Symmetry2C, data = diamonds,subset = index)))
}
set.seed(1)
boot.fn(diamonds,sample(400,400, replace = T))
boot.out <- boot(diamonds, boot.fn, 5000)
plot(boot.out)
```
# We can see a roughly normal distribution with the model which means the model is good and there is minimal overfitting.





<!-- ### Junk Transformation of Colour variable so that the groups of letter all correspond to the correct grouping -->
<!-- ```{r} -->
<!-- for(v in 1:440){ -->
<!-- if(diamonds$Colour[v] == "D"|| diamonds$Colour[v] == "E"|| diamonds$Colour[v] == "F") -->
<!-- { -->
<!--   diamonds$Colour[v] = 1 -->
<!-- } else if(diamonds$Colour[v] == "G"|| diamonds$Colour[v] == "H"|| diamonds$Colour[v] == "I") { -->
<!--   diamonds$Colour[v] = 2 -->

<!-- } else if(diamonds$Colour[v] == "J"|| diamonds$Colour[v] == "K") { -->
<!--   diamonds$Colour[v] = 3 -->

<!-- } else if(diamonds$Colour[v] == "L"|| diamonds$Colour[v] == "M"|| diamonds$Colour[v] == "N") { -->
<!--   diamonds$Colour[v] = 4 -->

<!-- } else if(diamonds$Colour[v] == "O"|| diamonds$Colour[v] == "P"|| diamonds$Colour[v] == "Q"|| diamonds$Colour[v] == "R"|| diamonds$Colour[v] == "S") { -->
<!--   diamonds$Colour[v] = 5 -->

<!-- } else  { -->
<!--   diamonds$Colour[v] = 6 -->

<!-- }  -->

<!-- } -->
<!-- ``` -->






<!-- #### Junk -->
<!-- ```{r} -->
<!-- diamonds <- read_excel("C:/Users/prana/Downloads/Professor_Proposes_Data.xlsx") -->
<!-- attach(diamonds) -->
<!-- diamonds$Colour[1] = "I" -->
<!-- if(diamonds$Colour[5] == "I" || diamonds$Colour[5] == "J") -->
<!-- { -->
<!--   diamonds$Colour[5] = 1 -->
<!-- } -->
<!-- diamonds$Colour[5] -->
<!-- #lm.fit <- lm(Price ~ Cut, data = diamonds) -->
<!-- #lm.fit -->



<!-- diamond_col_name <- colnames(diamonds) -->

<!-- #for (d in 1:9) { -->
<!--  # if(d !==8){ -->
<!--   #  lmfao[d] <- diamond_col_name[d] -->
<!--    # ggplot(diamonds, aes(x = diamond_col_name[d])) + -->
<!--              #geom_bar() -->
<!--   #} -->
<!-- #} -->

<!-- ``` -->



